Latent Variable Clustering for Classification of Events in the AT-TPC

### 2020-06-12 doc/Manuscript/article.tex
Updates from Overleaf

\begin{abstract}
% Modified by DB (6/8/2020)
This article presents novel applications of machine learning methods to the problem of event separation in an active target detector, the Active-Target Time Projection Chamber (AT-TPC) \cite{Bradt2017}. 
The overarching goal is to identify classes of events in the early stages of data processing, before 
%Machine learning presents an interesting avenue for researchers operating the AT-TPC, as traditional analysis methods of AT-TPC data are computationally expensive and have to fit all tracks against the event type of interest. The latter presents a considerable challenge when either the space of reactions is not known prior to the analysis, or the event type of interest is not the most abundant. 
The application of unsupervised clustering  algorithms to the analysis of two-dimensional projections of particle tracks from a resonant proton scattering experiment on $^{46}$Ar is introduced. We explore the performance of autoencoder neural networks and a pre-trained VGG16 \cite{Simonyan2014} convolutional neural network.
%on two tasks: a semi-supervised classification task and the unsupervised clustering of particle tracks. 
%On the semi-supervised task, we find that a logistic regression classifier trained on small labelled subsets of the latent space of these models perform very well, both on simulated and real data. On the clustering task, 
We find that a $K$-means algorithm applied to the simulated data in the VGG16 latent space forms almost perfect clusters. Additionally, the VGG16+K-means approach finds high purity clusters of proton events for real experimental data. We also explore the application of neural networks to clustering by implementing a mixture of autoencoders algorithm, with improved results on real data.

>>>>

\begin{abstract}
% Modified by DB (6/8/2020)
This article presents novel applications of machine learning methods to the problem of event separation in an active target detector, the Active-Target Time Projection Chamber (AT-TPC) \cite{Bradt2017}. 
The overarching goal is to identify classes of events in the early stages of the data analysis, thereby improving its efficiency by limiting the computationally expensive processing to events of interest only, which are often not the most abundant.
%Machine learning presents an interesting avenue for researchers operating the AT-TPC, as traditional analysis methods of AT-TPC data are computationally expensive and have to fit all tracks against the event type of interest. The latter presents a considerable challenge when either the space of reactions is not known prior to the analysis, or the event type of interest is not the most abundant. 
The application of unsupervised clustering  algorithms to the analysis of two-dimensional projections of particle tracks from a resonant proton scattering experiment on $^{46}$Ar is introduced. We explore the performance of autoencoder neural networks and a pre-trained VGG16 \cite{Simonyan2014} convolutional neural network.
%on two tasks: a semi-supervised classification task and the unsupervised clustering of particle tracks. 
%On the semi-supervised task, we find that a logistic regression classifier trained on small labelled subsets of the latent space of these models perform very well, both on simulated and real data. On the clustering task, 
We find that a $K$-means algorithm applied to the simulated data in the VGG16 latent space forms almost perfect clusters. Additionally, the VGG16+K-means approach finds high purity clusters of proton events for real experimental data. We also explore the application of neural networks to clustering by implementing a mixture of autoencoders algorithm, with improved results on real data.
